---
title: "Survival outcomes"
output: rmarkdown::html_vignette
description: >
  Bayesian DCA for survival (or time-to-event) outcomes.
vignette: >
  %\VignetteIndexEntry{Survival outcomes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.height = 7,
  dpi = 300
)
```

To illustrate Bayesian DCA on survival data, we load the example dataset `dca_survival_data` shown below.

```{r}
library(bayesDCA)
head(dca_survival_data)
```

The dataset contains the time-to-event `outcomes` (a `survival::Surv` object), the predictions from a prognostic model, and the results from a binary prognostic test. The time horizon for the event prediction is one time units (e.g., year), so we set `prediction_time = 1`. We also set `chains = 1` to speed up MCMC sampling with [Stan](https://mc-stan.org/) (in practice you should use at least `chains = 4`, maybe with `cores = 4` for speed as well).

```{r, fig.height=4.5}
fit <- dca_surv(dca_survival_data, prediction_time = 1, iter = 1000, chains = 1)
plot(fit)
```

# Interrogating the output

We can do all kinds of output interrogation just like with binary outcomes.

## What is the best decision strategy?

```{r}
compare_dca(fit)
```

## Are the prediction model or the binary test useful at all?

```{r}
compare_dca(fit, type = "useful")
```

## Is the model better than the test?

```{r fig.width=14}
compare_dca(fit, 
            strategies = c("model_predictions", "binary_test"), 
            type = "pairwise")
```

## What is the price of uncertainty?

```{r}
compare_dca(fit, .evpi = TRUE)
```



